Dynamics of thin liquid film flows studied with the lattice Boltzmann method

### 2023-08-10 chapters/Conclusion.tex
All references polished, conclusion missing

To this end a novel approach for modelling the dynamics of thin liquid films has been developed, see Chap.~\ref{chapter:theory}-\ref{chapter:second_paper}.
That said thin film flows have been studied with other models but all of them have shortcomings (the same is of course true for the model presented here).
The key question that has not been asked in Chaps.~\ref{chapter:intro}-\ref{chapter:third_paper} is: Is there a need for another method, or are there issues that can not be overcome with present solvers? 
The second part of this question can be answered with a cautions \textit{no}.
Many reseachers invested a lot of effort in the development of methods and solvers, problems such as dewetting on patterned substrates can be studied with several other solvers.
The first part of the question is a subjective one, because what do we actually need to numerical approximate the thin film equation?
A well posed finite difference scheme is often enough for simple problems.
However when the problems get complex due to couplings such as fluid structure interactions or phase changes (evaporation) we would prefer to use a method that we understand which is trustworthy and has been thoroughly tested. 
I think this one of the strengths of this method, the transparent nature of the software. 
The code base is available under an open source license and thus is freely available to everyone interested.
The package approach, similar to e.g. Python, Matlab or R, makes it simple to share and collaborate with colleagues on numerical experiments.
Of course there is a long list of thin film problems that have been studied with numerical models which will not benefit from our approach.
Spin coating is one of these problems that are outside the capabilities of our method.

>>>>

To this end a novel approach for modelling the dynamics of thin liquid films has been developed, see Chap.~\ref{chapter:theory}-\ref{chapter:second_paper}.
That said thin film flows have been studied with other models but all of them have shortcomings (the same is of course true for the model presented here).
% More thinking here!
The key question that has not been asked in Chaps.~\ref{chapter:intro}-\ref{chapter:third_paper} is: Is there a need for another method, or are there issues that can not be overcome with present solvers? 
The second part of this question can be answered with a cautions \textit{no}.
Many researchers invested a lot of effort in the development of methods and solvers, problems such as dewetting on patterned substrates can be studied with several other solvers.
The first part of the question is a subjective one, because what do we actually need to numerical approximate the thin film equation?
A well posed finite difference scheme is often enough for simple problems.
However when the problems get complex due to couplings such as fluid structure interactions or phase changes (evaporation) we would prefer to use a method that we understand which is trustworthy and has been thoroughly tested. 
I think this one of the strengths of this method, the transparent nature of the software. 
The code base is available under an open source license and thus is freely available to everyone interested.
The package approach, similar to e.g. Python, Matlab or R, makes it simple to share and collaborate with colleagues on numerical experiments.
Of course there is a long list of thin film problems that have been studied with numerical models which will not benefit from our approach.
Spin coating is one of these problems that are outside the capabilities of our method.

### 2023-08-10 chapters/Introduction.tex
All references polished, conclusion missing

Of course, by the end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g. the spectrum of capillary waves~\cite{vrijRuptureThinLiquid1968}.
While the spectrum was quickly connected with thermal fluctuations it wasn't an easy challenge to find a consistent theoretical description.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
The entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{landauFluidMechanicsLandau2013}.
While the mass flux is conserved at all times, heat transfer can violate Fouriers law at molecular scales~\cite{bellNumericalMethodsStochastic2007}.
Landau and Lifshitz derived a consistent stochastic stress term that accounts for these fluctuations and they added it to the Navier-Stokes equation.
Their stochastic term has zero mean, therefore the fluctuations do not create or destroy energy, and covariance proportional to thermal energy and viscosity. 
The stochastic thin film equation is then, again, derived from the LLNS equation by integration along the minor dimension, usually considered the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative Gaussian term~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, fetzerThermalNoiseInfluences2007, zhangNanoscaleThinfilmFlows2020, zhangMolecularSimulationThin2019, nesicFullyNonlinearDynamics2015}. 
This idea will be revisited in Chapter~\ref{chapter:second_paper}.
Around the same time Davidovitch et al., came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{davidovitchSpreadingViscousFluid2005}.
The addition of a Gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{grunThinFilmFlowInfluenced2006, fetzerThermalNoiseInfluences2007}.
Becker et al. pointed out that the classical thin film equation comes short in the prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations do lead to decreased rupture times~\cite{duran-olivenciaInstabilityRuptureFluctuations2019, shahThermalFluctuationsCapillary2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhangNonNewtonianEffectsLubricant2005, myersApplicationNonNewtonianModels2005}. 
Non-Newtonian liquids have a different response to stress than e.g. water. 
A short discussion about shear thickening, thinning liquids, such as toothpaste, follows in Chapter~\ref{chapter:theory}.
Other than the liquids' rheology it is possible to include arbitrary concentration fields and introduce Marangoni-like flows~\cite{sultanEvaporationThinFilm2005, hermansLungSurfactantsDifferent2015, surSteadyProfileFingeringFlows2004}.
Thermal fluctuations pose another addition of complexity as shown above~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, davidovitchSpreadingViscousFluid2005, zitzLatticeBoltzmannSimulations2021}.

On the other hand the substrate can be e.g. chemically or topographically patterned.
A modified substrate can lead to a very different understanding of wetting behaviour as shown in Refs.~\cite{cassieWettabilityPorousSurfaces1944, whymanRigorousDerivationYoung2008}.
Several research programs revolved around wetting, capillarity and the thin film equation. 
Within Europa the German Research Foundation (DFG) funded the priority program 1052: ``Wetting and Structure Formation'' from 1998 to 2004.
Karin Jacobs as project leader, one particularly interesting outcome was the fundamental work about dewetting morphology which was carried out by Becker et al.~\cite{beckerComplexDewettingScenarios2003}. 
Recently, in 2019, the DFG funded another priority program on wetting, called priority program 2171: ``\textit{Dynamic wetting of flexible, adaptive and switchable surfaces}''. 
The goal of this program is to enlarge the understanding concerning the dynamics of thin liquid films on complex substrates.
Complex substrates can for example be substrates with grafted polymer brushes~\cite{thieleGradientDynamicsModel2020}.
These brushes can swell somewhat similar to a sponge, allowing for special wetting and dewetting behaviour on such substrates. 
One treatment the substrate to generate a self-cleaning superhydrophobic surface similar to the lotus leaf~\cite{liAdaptationStyreneAcrylic2021, wongMicrodropletContaminantsWhen2020}.
Picking up the example of the lotus leaf, it is not necessary that the substrate is stiff.
In fact most organic structures are not as stiff as a silica substrate (or the glass of a window), therefore wetting of soft materials is as well considered in this priority program~\cite{andreottiStaticsDynamicsSoft2020, alandUnifiedNumericalModel2021, chenShortTimeWetting2011}.
More so with special surface treatment it is possible to control the substrates surface chemistry~\cite{xinReversiblySwitchableWettability2010, wangPhotoresponsiveSurfacesControllable2007}.
These stimuli responsive substrates are collected under the umbrella of the term \textit{switchable substrates}.
Allowing for the dynamic control of the substrates wettability with e.g. light sources~\cite{ichimuraLightDrivenMotionLiquids2000, sekiWideArrayPhotoinduced2018}.
\begin{figure}
    \centering
    \includegraphics[width=0.88\textwidth]{graphics/no_v_lig.pdf}
    \caption{(a) Equilibrium state of a dewetting thin film on a patterned substrate, i.e. droplets inside of the contact angle minima. 
    The color corresponds to the contact angle. 
    Yellow shows a high and blue a low contact angle, ranging from 10$^{\circ}$ to 30$^{\circ}$. 
    (b) Similar to the left panel a dewetting thin film on a ``switchable'' substrate. 
    Due to the function that controls the contact angle field liquid rivulets are formed.}
    \label{fig:morph_transition}
\end{figure}
This opens up a new avenue for microfluidic devices.
In Chapter~\ref{chapter:third_paper} we discuss the influence a switchable wettability can have on the dewetting of a thin film.
In fact the switching can induce morphological transitions as depict in Fig.~\ref{fig:morph_transition}.

@@

Using this principle creates a testing suite that (hopefully) ensures that functionality is not broken if changes to the source code are made.
For mature software projects, e.g. OpenFOAM~\cite{jasakOpenFOAMLibraryComplex, jasakOpenFOAMOpenSource2009, chenOpenFOAMComputationalFluid2014}, these suits can and should be even integrated into the continuous development/continuous integration (CD/CI) step and therefore be automated.
Often however when a student starts a new software project, functionality or performance is the main point of interest.
During the development step tests and even more important documentation are left aside.
Ultimately this does not only open the gate for various bugs but to some extent it also renders the software useless for other researchers.
Needless to say, my very own CFD solver, see Chapter~\ref{chapter:fourth_paper}, generated correct results due to the cancellation of two bugs.
Defining a test suite from the beginning of the project would have spared me valuable time and effort.

On the other hand, research became more than ever dependent on computational resources.
Clearly the main topic of this thesis, CFD is a compute heavy subdomain of fluid dynamics. 
However, recent advances in artificial intelligence show a rising trend to their applicability for industrial and scientific problems~\cite{acemogluArtificialIntelligenceAutomation2018, beintemaControllingRayleighBenard2020}.
These advances comes at a price, leaving aside the discussion if artificial intelligence is of value at all.
One of the main pillars that generates trust in scientific results is their reproducibility.
Independent of the number of trials a stone, when flung up in the air, will fall to the ground following a well defined path\footnote{Given the stone is not flung with earth's escape velocity, $v\approx 11.2$km/s}.
Generating huge amounts of data without labels and explanation makes it by definition impossible to verify any given finding or result from that experiment.
Mislabeling data is not something new, it always happened from time to time with e.g. laboratory journals.
However, exa-scaling computing pushes this problem to new severity.
Now or never is the time to install protocols. 
Guides on how to deal with scientific software such that data can be reproduced or even more important reused.
While this may not be of interest for most readers, to me it is an essential point.
That said, Chapter~\ref{chapter:fourth_paper} describes the software developed and used for this thesis. 
One can find the open source software repository that hosts the code, tests and documentation.
Upon pull or merge requests an automated testing suite performs tests to ensure no bugs break necessary functionality.
Furthermore by definition of version control it is possible to reproduce all results generated with that software, called \textit{Swalbe.jl}.

\section{Outline}
\label{section:outline}
Following this short introduction are chapters on theory, the numerical method, results and finally a conclusion.
Starting with the next Chapter~\ref{chapter:theory} a detailed explanation of the theoretical ideas is presented.
Thus the starting point will be the equation of motion of a fluid, the Navier-Stokes equation.
Using real world observations and strong assumptions the Navier-Stokes equation can be reduced to the Saint-Venant or shallow water equation~\cite{bTheorieMouvementNonpermanent1871}.
Similar arguments also play a critical role in the derivation of the thin film equation.
However, these two systems describe different physical phenomena.
Still, as will be shown, this is a direct consequence of \textit{long-scale}, \textit{long-wave}  phenomena~\cite{oronLongscaleEvolutionThin1997}.
Chapter~\ref{chapter:theory} will then end with a short section of differences and intersections between these two theories.

In Chapter~\ref{chapter:method} numerical frameworks for solving differential equations are presented.
Starting with a short overview on different methods to be used to solve the thin film problem.
Specifically, the concept of discrete differentiation is introduced as a consequence of a discrete computational domain. 
One strategy to avoid differentiation is to use spectral methods which operate via Fourier transformation. 
Lastly one method that offers a lot of value for nanometric thin films is molecular dynamics (MD).
As it turns out different length and time scales can be a pitfall for either of these methods.
In between the macro (classical Navier-Stokes solvers) and micro (MD) scale is the lattice Boltzmann method (LBM), a so-called meso-scale approach.
The underlying differential equation that is solved is the Boltzmann equation. 
LBM is therefore not another discretization scheme for the Navier-Stokes equation. 
However, as will be shown the Navier-Stokes equation can be approximated using the lattice Boltzmann method~\cite{enskogKinetischeTheorieVorgange1917, chapmanMathematicalTheoryNonuniform1990, chenLatticeBoltzmannMethod1998, krugerLatticeBoltzmannMethod2017}.  
Not only is it possible to approximate the Navier-Stokes but also the shallow water equations, of course with a different set of constraints~\cite{salmonLatticeBoltzmannMethod1999, zhouLatticeBoltzmannMethods2004, vanthangStudy1DLattice2010, dellarNonhydrodynamicModesPriori2002}.

Following that argumentation is the published article for the \textit{Journal of Open Source Software}. 
Highlighting the implementation and development of the lattice Boltzmann solver called \textbf{Swalbe} (\textbf{s}hallow \textbf{wa}ter \textbf{l}attice \textbf{B}oltzmann solv\textbf{e}r) in Chapter~\ref{chapter:fourth_paper}.
Key aspects are the automated continuous integration (CI) with a test suite and web-hosted documentation, as motivated in Sec.~\ref{section:statement_software}.
The Julia package allows for either fast iterative model development in two dimensions or large scale simulations in three dimensions with GPU acceleration.
It allows for prototyping and testing which is fast and easy to use, but also a framework for large three dimensional simulations with minor to no code changes.

% I am here
Use cases and a derivation of the model can be found in Chapter~\ref{chapter:first_paper}.
In this chapter the mandatory modelling assumptions are introduced that allow to match the shallow water system with the thin film equation.
It further highlights which modifications are made to the shallow water lattice Boltzmann algortihm. 
With some emphasis on the numerical implementation of e.g. the computation of gradients and the Laplacian in agreement with Ref~\cite{junkDiscretizationsIncompressibleNavier2000, thampiIsotropicDiscreteLaplacian2013}. 
After the derivation, numerical experiments are displayed to validate the model. 
Among those are the Rayleigh-Taylor instability, an instability that occurs when a heavy fluid is on top of light fluid. 
Spreading dynamics is tested against the Cox-Voinov and Tanner's laws. 
The linear relation between Bond and Capillary number is validated with the sliding of a droplet.
And for completeness a discussion on the performance of the algorithm and applicability for accelerated computing at the end of the chapter.

@@

% I am here
Dynamics during dewetting can either be generated due to forces, as shown in Chapters~\ref{chapter:first_paper}-\ref{chapter:second_paper} or it can be induced due to time dependent potentials. 
Giving the recent interest in dynamics of simple liquids on complex substrates and the dynamic wetting of flexible, adaptive and switchable surfaces, Chapter~\ref{chapter:fourth_paper} is dedicated to the question: ``What happens during dewetting with a spatio-temporal evolving wettability gradient?''
To study this problem a series of three dimensional simulations is performed with varying spatial as well as temporal evolution of the wettability.
On the one hand the wettability dynamics has a small but measurable stabilizing effect on the spinodal dewetting of the thin film leading to a net increase in rupture times if the dynamic wetting is switched on.
On the other hand after the rupture of the film a clear morphological transition is observable.
Having only a static wettability gradient results in stationary droplets in regions of high wettability.
Making the wettability gradient dynamic can lead to metastable rivulet states.
These rivulets can be explained using a simple model that balances the capillary with the wettability wave velocity, introducing a cutoff parameter ($\Gamma > 1$) for which rivulets are to be observed.

>>>>

Of course, by the end of the 20$^{\text{th}}$ century there were already some yet unexplained experimental findings associated to the thin film equation, e.g. the spectrum of capillary waves~\cite{vrijRuptureThinLiquid1968}.
While the spectrum was quickly connected with thermal fluctuations it wasn't an easy challenge to find a consistent theoretical description.
From a theoretical point of view one has to account for the stochastic nature of thermal fluctuations.
The entry point for this problem is not the Navier-Stokes equation but the Landau-Lifshitz Navier-Stokes (LLNS) equation~\cite{landauFluidMechanicsLandau2013}.
While the mass flux is conserved at all times, heat transfer can violate Fouriers law at molecular scales~\cite{bellNumericalMethodsStochastic2007}.
Landau and Lifshitz derived a consistent stochastic stress term that accounts for these fluctuations and they added it to the Navier-Stokes equation.
Their stochastic term has zero mean, therefore the fluctuations do not create or destroy energy, and covariance proportional to thermal energy and viscosity. 
The stochastic thin film equation is then, again, derived from the LLNS equation by integration along the minor dimension, usually considered the vertical dimension.
This step however poses a hard problem, because the stochastic stress tensor in principle depends on the vertical dimension.
Grün and coworkers found an elegant way to circumvent the problem with a single multiplicative Gaussian term~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, fetzerThermalNoiseInfluences2007, zhangNanoscaleThinfilmFlows2020, zhangMolecularSimulationThin2019, nesicFullyNonlinearDynamics2015}. 
This idea will be revisited in Chap.~\ref{chapter:second_paper}.
Around the same time Davidovitch et al., came up with a similar form for the stochastic thin film equation, studying the spreading of a droplet~\cite{davidovitchSpreadingViscousFluid2005}.
The addition of a Gaussian noise term made it therefore possible to reproduce the observed capillary wave spectrum and explain the difference in spreading behaviour.
A further impact of this term was the reduction of rupture times for spinodal dewetting thin films~\cite{grunThinFilmFlowInfluenced2006, fetzerThermalNoiseInfluences2007}.
Becker et al. pointed out that the classical thin film equation comes short in the prediction of the experimentally measured rupture times~\cite{beckerComplexDewettingScenarios2003}.
In fact the rupture time is overestimated by the classical thin film equation, adding however thermal fluctuations do lead to decreased rupture times~\cite{duran-olivenciaInstabilityRuptureFluctuations2019, shahThermalFluctuationsCapillary2019}.

Additions of complex dynamics on top of the partial differential equation that defines the thin film problems is not limited to thermal fluctuations.
On the one hand it is possible to add degrees of freedom to the fluid. 
For example one can study the effect of non-Newtonian shear behaviour~\cite{zhangNonNewtonianEffectsLubricant2005, myersApplicationNonNewtonianModels2005}. 
Non-Newtonian liquids have a different response to stress than e.g. water. 
A short discussion about shear thickening, thinning liquids, such as toothpaste, follows in Chap.~\ref{chapter:theory}.
Other than the liquids' rheology it is possible to include arbitrary concentration fields and introduce Marangoni-like flows~\cite{sultanEvaporationThinFilm2005, hermansLungSurfactantsDifferent2015, surSteadyProfileFingeringFlows2004}.
Thermal fluctuations pose another addition of complexity as shown above~\cite{grunThinFilmFlowInfluenced2006, meckeThermalFluctuationsThin2005, davidovitchSpreadingViscousFluid2005, zitzLatticeBoltzmannSimulations2021}.

On the other hand the substrate can be e.g. chemically or topographically patterned.
A modified substrate can lead to a very different understanding of wetting behaviour as shown in Refs.~\cite{cassieWettabilityPorousSurfaces1944, whymanRigorousDerivationYoung2008}.
Several research programs revolved around wetting, capillarity and the thin film equation. 
Within Europa the German Research Foundation (DFG) funded the priority program 1052: ``Wetting and Structure Formation'' from 1998 to 2004.
Karin Jacobs as project leader, one particularly interesting outcome was the fundamental work about dewetting morphology which was carried out by Becker et al.~\cite{beckerComplexDewettingScenarios2003}. 
Recently, in 2019, the DFG funded another priority program on wetting, called priority program 2171: ``\textit{Dynamic wetting of flexible, adaptive and switchable surfaces}''. 
The goal of this program is to enlarge the understanding concerning the dynamics of thin liquid films on complex substrates.
Complex substrates can for example be substrates with grafted polymer brushes~\cite{thieleGradientDynamicsModel2020}.
These brushes can swell somewhat similar to a sponge, allowing for special wetting and dewetting behaviour on such substrates. 
One treatment the substrate to generate a self-cleaning superhydrophobic surface similar to the lotus leaf~\cite{liAdaptationStyreneAcrylic2021, wongMicrodropletContaminantsWhen2020}.
Picking up the example of the lotus leaf, it is not necessary that the substrate is stiff.
In fact most organic structures are not as stiff as a silica substrate (or the glass of a window), therefore wetting of soft materials is as well considered in this priority program~\cite{andreottiStaticsDynamicsSoft2020, alandUnifiedNumericalModel2021, chenShortTimeWetting2011}.
More so with special surface treatment it is possible to control the substrates surface chemistry~\cite{xinReversiblySwitchableWettability2010, wangPhotoresponsiveSurfacesControllable2007}.
These stimuli responsive substrates are collected under the umbrella of the term \textit{switchable substrates}.
Allowing for the dynamic control of the substrates wettability with e.g. light sources~\cite{ichimuraLightDrivenMotionLiquids2000, sekiWideArrayPhotoinduced2018}.
\begin{figure}
    \centering
    \includegraphics[width=0.88\textwidth]{graphics/no_v_lig.pdf}
    \caption{(a) Equilibrium state of a dewetting thin film on a patterned substrate, i.e. droplets inside of the contact angle minima. 
    The color corresponds to the contact angle. 
    Yellow shows a high and blue a low contact angle, ranging from 10$^{\circ}$ to 30$^{\circ}$. 
    (b) Similar to the left panel a dewetting thin film on a ``switchable'' substrate. 
    Due to the function that controls the contact angle field liquid rivulets are formed.}
    \label{fig:morph_transition}
\end{figure}
This opens up a new avenue for microfluidic devices.
In Chap.~\ref{chapter:third_paper} we discuss the influence a switchable wettability can have on the dewetting of a thin film.
In fact the switching can induce morphological transitions as depict in Fig.~\ref{fig:morph_transition}.

@@

Using this principle creates a testing suite that (hopefully) ensures that functionality is not broken if changes to the source code are made.
For mature software projects, e.g. OpenFOAM~\cite{jasakOpenFOAMLibraryComplex, jasakOpenFOAMOpenSource2009, chenOpenFOAMComputationalFluid2014}, these suits can and should be even integrated into the continuous development/continuous integration (CD/CI) step and therefore be automated.
Often however when a student starts a new software project, functionality or performance is the main point of interest.
During the development step tests and even more important documentation are left aside.
Ultimately this does not only open the gate for various bugs but to some extent it also renders the software useless for other researchers.
Needless to say, my very own CFD solver, see Chap.~\ref{chapter:fourth_paper}, generated correct results due to the cancellation of two bugs.
Defining a test suite from the beginning of the project would have spared me valuable time and effort.

On the other hand, research became more than ever dependent on computational resources.
Clearly the main topic of this thesis, CFD is a compute heavy subdomain of fluid dynamics. 
However, recent advances in artificial intelligence show a rising trend to their applicability for industrial and scientific problems~\cite{acemogluArtificialIntelligenceAutomation2018, beintemaControllingRayleighBenard2020}.
These advances comes at a price, leaving aside the discussion if artificial intelligence is of value at all.
One of the main pillars that generates trust in scientific results is their reproducibility.
Independent of the number of trials a stone, when flung up in the air, will fall to the ground following a well defined path\footnote{Given the stone is not flung with earth's escape velocity, $v\approx 11.2$km/s}.
Generating huge amounts of data without labels and explanation makes it by definition impossible to verify any given finding or result from that experiment.
Mislabeling data is not something new, it always happened from time to time with e.g. laboratory journals.
However, exa-scaling computing pushes this problem to new severity.
Now or never is the time to install protocols. 
Guides on how to deal with scientific software such that data can be reproduced or even more important reused.
While this may not be of interest for most readers, to me it is an essential point.
That said, Chap.~\ref{chapter:fourth_paper} describes the software developed and used for this thesis. 
One can find the open source software repository that hosts the code, tests and documentation.
Upon pull or merge requests an automated testing suite performs tests to ensure no bugs break necessary functionality.
Furthermore by definition of version control it is possible to reproduce all results generated with that software, called \textit{Swalbe.jl}.

\section{Outline}
\label{section:outline}
Following this short introduction are chapters on theory, the numerical method, results and finally a conclusion.
Starting with the next Chap.~\ref{chapter:theory} a detailed explanation of the theoretical ideas is presented.
Thus the starting point will be the equation of motion of a fluid, the Navier-Stokes equation.
Using real world observations and strong assumptions the Navier-Stokes equation can be reduced to the Saint-Venant or shallow water equation~\cite{bTheorieMouvementNonpermanent1871}.
Similar arguments also play a critical role in the derivation of the thin film equation.
However, these two systems describe different physical phenomena.
Still, as will be shown, this is a direct consequence of \textit{long-scale}, \textit{long-wave}  phenomena~\cite{oronLongscaleEvolutionThin1997}.
Chap.~\ref{chapter:theory} will then end with a short section of differences and intersections between these two theories.

In Chap.~\ref{chapter:method} numerical frameworks for solving differential equations are presented.
Starting with a short overview on different methods to be used to solve the thin film problem.
Specifically, the concept of discrete differentiation is introduced as a consequence of a discrete computational domain. 
One strategy to avoid differentiation is to use spectral methods which operate via Fourier transformation. 
Lastly one method that offers a lot of value for nanometric thin films is molecular dynamics (MD).
As it turns out different length and time scales can be a pitfall for either of these methods.
In between the macro (classical Navier-Stokes solvers) and micro (MD) scale is the lattice Boltzmann method (LBM), a so-called meso-scale approach.
The underlying differential equation that is solved is the Boltzmann equation. 
LBM is therefore not another discretization scheme for the Navier-Stokes equation. 
However, as will be shown the Navier-Stokes equation can be approximated using the lattice Boltzmann method~\cite{enskogKinetischeTheorieVorgange1917, chapmanMathematicalTheoryNonuniform1990, chenLatticeBoltzmannMethod1998, krugerLatticeBoltzmannMethod2017}.  
Not only is it possible to approximate the Navier-Stokes but also the shallow water equations, of course with a different set of constraints~\cite{salmonLatticeBoltzmannMethod1999, zhouLatticeBoltzmannMethods2004, vanthangStudy1DLattice2010, dellarNonhydrodynamicModesPriori2002}.

Following that argumentation is the published article for the \textit{Journal of Open Source Software}. 
Highlighting the implementation and development of the lattice Boltzmann solver called \textbf{Swalbe} (\textbf{s}hallow \textbf{wa}ter \textbf{l}attice \textbf{B}oltzmann solv\textbf{e}r) in Chap.~\ref{chapter:fourth_paper}.
Key aspects are the automated continuous integration (CI) with a test suite and web-hosted documentation, as motivated in Sec.~\ref{section:statement_software}.
The Julia package allows for either fast iterative model development in two dimensions or large scale simulations in three dimensions with GPU acceleration.
It allows for prototyping and testing which is fast and easy to use, but also a framework for large three dimensional simulations with minor to no code changes.

% I am here
Use cases and a derivation of the model can be found in Chap.~\ref{chapter:first_paper}.
In this chapter the mandatory modelling assumptions are introduced that allow to match the shallow water system with the thin film equation.
It further highlights which modifications are made to the shallow water lattice Boltzmann algortihm. 
With some emphasis on the numerical implementation of e.g. the computation of gradients and the Laplacian in agreement with Ref~\cite{junkDiscretizationsIncompressibleNavier2000, thampiIsotropicDiscreteLaplacian2013}. 
After the derivation, numerical experiments are displayed to validate the model. 
Among those are the Rayleigh-Taylor instability, an instability that occurs when a heavy fluid is on top of light fluid. 
Spreading dynamics is tested against the Cox-Voinov and Tanner's laws. 
The linear relation between Bond and Capillary number is validated with the sliding of a droplet.
And for completeness a discussion on the performance of the algorithm and applicability for accelerated computing at the end of the chapter.

@@

% I am here
Dynamics during dewetting can either be generated due to forces, as shown in Chaps.~\ref{chapter:first_paper}-\ref{chapter:second_paper} or it can be induced due to time dependent potentials. 
Giving the recent interest in dynamics of simple liquids on complex substrates and the dynamic wetting of flexible, adaptive and switchable surfaces, Chap.~\ref{chapter:fourth_paper} is dedicated to the question: ``What happens during dewetting with a spatio-temporal evolving wettability gradient?''
To study this problem a series of three dimensional simulations is performed with varying spatial as well as temporal evolution of the wettability.
On the one hand the wettability dynamics has a small but measurable stabilizing effect on the spinodal dewetting of the thin film leading to a net increase in rupture times if the dynamic wetting is switched on.
On the other hand after the rupture of the film a clear morphological transition is observable.
Having only a static wettability gradient results in stationary droplets in regions of high wettability.
Making the wettability gradient dynamic can lead to metastable rivulet states.
These rivulets can be explained using a simple model that balances the capillary with the wettability wave velocity, introducing a cutoff parameter ($\Gamma > 1$) for which rivulets are to be observed.
